---
title: "Quality Control of scRNAseq DESCRIPTION"
author:
  - Chao-Chung Kuo, Genomics Facility, ckuo@ukaachen.de
  - Lin Gan, Genomics Facility, lgan@ukaachen.de
  - Mohamed Mabrouk, Genomics Facility, mmabrouk@ukaachen.de
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_fold: hide
    df_print: paged
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: false
    theme: lumen
---

```{r, echo=FALSE, results="hide", warning=FALSE, message=FALSE}
Sys.setenv(PATH = paste(Sys.getenv("PATH"), "/opt/miniconda3/envs/R4.3.2/bin", sep = ":"))
Sys.setenv(RENV_PATHS_ROOT = "/data/shared_env/renv/")
options(repos = BiocManager::repositories())

library(Seurat) # Seurat V5 is recommended
library(stringr)
library(ggplot2)
library(knitr)
library(gridExtra)
library(scuttle)
library(kableExtra)

echo_param <- FALSE
```

```{r, echo=echo_param, results='hide', warning=FALSE, message=FALSE}
sample_dir <- "SAMPLE_DIR"
file_TAG <- "FILE_TAG"

source("scRNAseq_QC_settings.RData")
nfeatures_low <- NULL
nfeatures_high <- NULL

ncounts_low <- NULL
ncounts_high <- NULL

percent_mito_low <- NULL
percent_mito_high <- NULL
```

```{r, echo=echo_param, results='hide', warning=FALSE, message=FALSE}
readers_factory <- list(
  "10X" = Seurat::Read10X,
  "ParseBio" = Seurat::ReadParseBio,
  "10X_h5" = Seurat::Read10X_h5
)

parser <- readers_factory[[technology]]

if (!is.null(parser)) {
  obj <- parser(INPUT_DIR)
} else {
  stop("Only '10X', '10X_h5', 'ParseBio' are supported! ")
}
```

```{r, echo=echo_param, results='hide', warning=FALSE, message=FALSE}
seurat_obj <- CreateSeuratObject(obj, min.cells = min_cells, min.features = min_features, project = basename(INPUT_DIR))
seurat_obj[["percent.mito"]] <- PercentageFeatureSet(seurat_obj, pattern = "^MT-")
seurat_obj[["percent.ribo"]] <- PercentageFeatureSet(seurat_obj, pattern = "^RPS|^RPL")
```

# Diagnostic plots {.tabset}
## Violin Plot
```{r violin_plot, echo=echo_param, results='asis', warning=FALSE, message=FALSE, dpi=300, fig.width=8, fig.height=6, fig.format='png'}
VlnPlot(seurat_obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mito", "percent.ribo"), ncol = 2)
```

```{r outlier calculation, echo= echo_param, results="hide", warning=FALSE, message=FALSE}
if (!is.null(ncounts_low) & !is.null(ncounts_high)) {
  count <- ncounts_low > seurat_obj$nCount_RNA | seurat_obj$nCount_RNA > ncounts_high
} else {
  count <- isOutlier(seurat_obj$nCount_RNA, nmads = nmads, type = "both")
}

if (!is.null(nfeatures_low) & !is.null(nfeatures_high)) {
  count <- nfeatures_low > seurat_obj$nFeature_RNA | seurat_obj$nFeature_RNA > nfeatures_high
} else {
  features <- isOutlier(seurat_obj$nFeature_RNA, nmads = nmads, type = "both")
}

if (!is.null(percent_mito_low) & !is.null(percent_mito_high)) {
  count <- percent_mito_low > seurat_obj$percent.mito | seurat_obj$percent.mito > percent_mito_high
} else {
  mito <- isOutlier(seurat_obj$percent.mito, nmads = nmads, type = "both")
}
```

## Histograms

```{r histograms, echo=echo_param, results='asis', warning=FALSE, message=FALSE, dpi=300, fig.width=8, fig.height=6, fig.format='png', }
histogram_seurat <- function(obj, variable, n_mad = 3, vlines = FALSE, annotate = FALSE, no_breaks = 30) {
  vec <- obj@meta.data[[variable]]
  mad_features <- mad(vec)
  median_features <- median(vec)
  upper <- median_features + n_mad * mad_features
  lower <- median_features - n_mad * mad_features

  upper_cells <- sum(vec > upper)
  lower_cells <- sum(vec < lower)
  inbetween <- sum(lower < vec & vec < upper)

  info <- hist(vec, breaks = no_breaks, plot = FALSE)
  y_pos <- mean(sort(info$counts, decreasing = TRUE)[1:3])



  fig <- ggplot(obj@meta.data, aes_string(x = variable)) +
    geom_histogram(fill = "#86BAFE", color = "#5DA3FD", bins = no_breaks) +
    theme_classic()
  if (vlines == TRUE) {
    fig <- fig + geom_vline(xintercept = upper, linetype = 2) +
      geom_vline(xintercept = lower, linetype = 2)
  }

  if (annotate == TRUE) {
    fig <- fig + annotate(geom = "text", label = upper_cells, x = (upper + max(info$breaks)) / 2, y = y_pos) +
      annotate(geom = "text", label = lower_cells, x = (lower + min(info$breaks)) / 2, y = y_pos) +
      annotate(geom = "text", label = inbetween, x = mean(vec), y = y_pos * 1.1) +
      ylab("count")
  }

  return(fig)
}

p1 <- histogram_seurat(seurat_obj, "nFeature_RNA", nmads, hist_vlines, hist_annotate, no_breaks)
p2 <- histogram_seurat(seurat_obj, "percent.mito", nmads, hist_vlines, hist_annotate, no_breaks)
p3 <- histogram_seurat(seurat_obj, "nCount_RNA", nmads, hist_vlines, hist_annotate, no_breaks)
p4 <- histogram_seurat(seurat_obj, "percent.ribo", nmads, hist_vlines, hist_annotate, no_breaks)


(p1 + p2) / (p3 + p4)
```

```{r, echo=echo_param, results='hide', warning=FALSE, message=FALSE}
seurat_obj@meta.data$features_outlier <- features
seurat_obj@meta.data$counts_outlier <- count
seurat_obj@meta.data$mito_outlier <- mito

mask <- count | features

if (subset_mito) {
  mask <- mask | mito
}

seurat_obj@meta.data$outlier <- mask
```

# Detection of low-quality cells {.tabset}

## Scatter Plot of Outliers

```{r scatter_plots, echo=echo_param, results='asis', warning=FALSE, message=FALSE, dpi=300, fig.width=10, fig.height=4, fig.format='png'}
plot1 <- ggplot(as.data.frame(seurat_obj@meta.data), aes(x = nCount_RNA, y = percent.mito, color = mito_outlier)) +
  geom_point(size = 1, alpha = 0.4) +
  theme_classic() +
  scale_color_brewer(palette = "Set2")

plot2 <- ggplot(as.data.frame(seurat_obj@meta.data), aes(x = nCount_RNA, y = nFeature_RNA, color = outlier)) +
  geom_point(size = 1, alpha = 0.4) +
  theme_classic() +
  scale_color_brewer(palette = "Set2")

plot1 + plot2
```



```{r default_analysis_pipeline_before, echo=echo_param, results='hide', warning=FALSE, message=FALSE}
seurat_obj <- seurat_obj %>%
  NormalizeData() %>%
  FindVariableFeatures() %>%
  ScaleData()
seurat_obj <- seurat_obj %>% RunPCA(features = VariableFeatures(object = seurat_obj))
seurat_obj <- seurat_obj %>%
  FindNeighbors(dims = 1:20) %>%
  FindClusters() %>%
  RunUMAP(dims = 1:20)
```

# Before removal of low-quality cells {.tabset}

## PCA

```{r dim_reductionplot, echo=echo_param, results='asis', warning=FALSE, message=FALSE, dpi=300, fig.width=4, fig.height=3, fig.format='png'}
DimPlot(seurat_obj, reduction = "pca", group.by = "orig.ident") +
  theme_classic() +
  theme(text = element_text(size = 10))
```

## UMAP

```{r dim_red_umap, echo=echo_param, results='asis', warning=FALSE, message=FALSE, dpi=300, fig.width=4, fig.height=3, fig.format='png'}
DimPlot(seurat_obj, reduction = "umap", group.by = "orig.ident") +
  theme_classic() +
  theme(text = element_text(size = 10))
```

# Confounders before removal of low-quality cells {.tabset}

## UMAP_nfeatures

```{r featureplot_nfeatures, echo=echo_param, results='asis', warning=FALSE, message=FALSE, dpi=300, fig.width=4, fig.height=3, fig.format='png'}
FeaturePlot(object = seurat_obj, features = "nFeature_RNA") +
  theme_classic() +
  theme(text = element_text(size = 10))
```

## UMAP_nCounts

```{r featureplot_ncounts, echo=echo_param, results='asis', warning=FALSE, message=FALSE, dpi=300, fig.width=4, fig.height=3, fig.format='png'}
FeaturePlot(object = seurat_obj, features = "nCount_RNA") +
  theme_classic() +
  theme(text = element_text(size = 10))
```

## UMAP_percent.mito

```{r featureplot_percent.mito, echo=echo_param, results='asis', warning=FALSE, message=FALSE, dpi=300, fig.width=4, fig.height=3, fig.format='png'}
FeaturePlot(object = seurat_obj, features = "percent.mito") +
  theme_classic() +
  theme(text = element_text(size = 10))
```


```{r subset_followed_by_analysis, echo=echo_param, results='hide', warning=FALSE, message=FALSE, eval= subset}
# Evaluation is controlled via the subset parameter in RMarkdown options
seurat_obj <- subset(seurat_obj, outlier == FALSE)
seurat_obj <- seurat_obj %>%
  NormalizeData() %>%
  FindVariableFeatures() %>%
  ScaleData()
seurat_obj <- seurat_obj %>% RunPCA(features = VariableFeatures(object = seurat_obj))
seurat_obj <- seurat_obj %>%
  FindNeighbors(dims = 1:20) %>%
  FindClusters() %>%
  RunUMAP(dims = 1:20)
```


`r if(subset){"# Results after removal of low-quality cells {.tabset}"}` `r if(subset){"## PCA"}`

```{r dim_reductionplot_after, eval=subset ,echo=echo_param, results='asis', warning=FALSE, message=FALSE, dpi=300, fig.width=4, fig.height=3, fig.format='png'}
DimPlot(seurat_obj, reduction = "pca", group.by = "orig.ident") +
  theme_classic() +
  theme(text = element_text(size = 10))
```

`r if(subset){"## UMAP"}`

```{r dim_red_umap_after, eval=subset, echo=echo_param, results='asis', warning=FALSE, message=FALSE, dpi=300, fig.width=4, fig.height=3, fig.format='png'}
DimPlot(seurat_obj, reduction = "umap", group.by = "orig.ident") +
  theme_classic() +
  theme(text = element_text(size = 10))
```

`r if(subset){"# Confounders after removal of low-quality cells {.tabset}"}` `r if(subset){"## UMAP_nfeatures"}`

```{r featureplot_nfeatures_after, eval=subset, echo=echo_param, results='asis', warning=FALSE, message=FALSE, dpi=300, fig.width=4, fig.height=3, fig.format='png'}
FeaturePlot(object = seurat_obj, features = "nFeature_RNA") +
  theme_classic() +
  theme(text = element_text(size = 10))
```

`r if(subset){"## UMAP_nCounts"}`

```{r featureplot_ncounts_after, eval=subset, echo=echo_param, results='asis', warning=FALSE, message=FALSE, dpi=300, fig.width=4, fig.height=3, fig.format='png'}
FeaturePlot(object = seurat_obj, features = "nCount_RNA") +
  theme_classic() +
  theme(text = element_text(size = 10))
```

`r if(subset){"## UMAP_percent.mito"}`

```{r featureplot_percent.mito_after, eval=subset, echo=echo_param, results='asis', warning=FALSE, message=FALSE, dpi=300, fig.width=4, fig.height=3, fig.format='png'}
FeaturePlot(object = seurat_obj, features = "percent.mito") +
  theme_classic() +
  theme(text = element_text(size = 10))
```

`r if(subset){"# {-}"}`

# R session information

```{r, echo=echo_param, results='markup', warning=FALSE, message=TRUE}
sessionInfo()
```
