---
title: "HTOs Demultiplex"
author: 
PROJECT_AUTHORS
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: false
    theme: lumen
    fig_width: 5  # Adjust the width as needed
    fig_height: 6
---

Reference: [Demultiplexing with hashtag oligos (HTOs)](https://satijalab.org/seurat/articles/hashing_vignette.html)


```{r init, echo=FALSE, results="hide", warning=FALSE, message=FALSE}
source("/etc/rstudio/Rprofile.site")
library(Seurat)
library(stringr)
library(ggplot2)
library(knitr)
library(gridExtra)

GEX_names <- list.files(path = "PROECCT_PROCESSING_PATH/results/cellranger/mtx_conversions/", 
                         pattern = "*_matrix.rds", full.names = FALSE)
GEX_names <- sub("_matrix.rds$", "", GEX_names)
GEX_names
Hashtag_names <- list.dirs(path = "PROECCT_PROJECT_PATH/citeseq_count/", full.names = FALSE, recursive = FALSE)
Hashtag_names

GEX_dir <- "PROECCT_PROCESSING_PATH/results/cellranger/mtx_conversions/"
Hashtag_dir <- "PROECCT_PROJECT_PATH/citeseq_count/"

demultiplex_GEX <- function(GEX_dir, GEX_name, Hashtag_dir, Hashtag_name, hashtag_labels, hashtag_index) {
  # Load GEX matrix
  GEX_matrix <- readRDS(paste0(GEX_dir, GEX_name, "_matrix.rds"))
  # Check the frequency of UMI information
  table(str_extract(colnames(GEX_matrix), "\\d+"))
  
  colnames(GEX_matrix) <- sub("-1$", "", colnames(GEX_matrix))
  # Load HTO matrix
  HTOs_matrix <- Read10X(paste0(Hashtag_dir, Hashtag_name, "/umi_count/"), gene.column=1)
  # cat(length(rownames(HTOs_matrix)))
  
  rownames(HTOs_matrix) <- sub("-.*", "", rownames(HTOs_matrix))
  rownames(HTOs_matrix)[hashtag_index] <- hashtag_labels
  HTOs_matrix <- HTOs_matrix[hashtag_index,]
  # Find overlapping cell barcodes
  # joint.bcs <- intersect(sub("-1$", "", colnames(GEX_matrix)), colnames(HTOs_matrix))
  joint.bcs <- intersect(colnames(GEX_matrix), colnames(HTOs_matrix))
  cell_num <- data.frame(Library=c("GEX library", "Hashtag library", "Joint cell barcodes"),
                         Cells=c(length(colnames(GEX_matrix)),
                                 length(colnames(HTOs_matrix)),
                                 length(joint.bcs)))
  
  # Subset GEX and HTO counts by joint cell barcodes
  GEX_matrix <- GEX_matrix[, joint.bcs]
  HTOs_matrix <- as.matrix(HTOs_matrix[, joint.bcs])
  # Confirm that the HTO have the correct names
  rownames(HTOs_matrix)
  
  # Setup Seurat object
  hashtag <- CreateSeuratObject(counts = Matrix::Matrix(GEX_matrix[["RNA"]]@counts, sparse = T))
  # Normalize RNA data with log normalization
  hashtag <- NormalizeData(hashtag)
  # Find and scale variable features
  hashtag <- FindVariableFeatures(hashtag, selection.method = "mean.var.plot")
  hashtag <- ScaleData(hashtag, features = VariableFeatures(hashtag))
  
  #Adding HTO data as an independent assay
  # Add HTO data as a new assay independent from RNA
  hashtag[["HTO"]] <- CreateAssayObject(counts = HTOs_matrix)
  # Normalize HTO data, here we use centered log-ratio (CLR) transformation
  hashtag <- NormalizeData(hashtag, assay = "HTO", normalization.method = "CLR")
  
  # Demultiplex cells based on HTO enrichment
  hashtag <- HTODemux(hashtag, assay = "HTO", positive.quantile = 0.99)
  
  hto_freq <- as.data.frame(table(hashtag@meta.data$HTO_maxID))
  hto_freq$Var1 <- as.character(hto_freq$Var1)
  colnames(hto_freq) <- c("Sample", "Cell number")
  
  print(knitr::kable(cell_num, caption = "Cell barcodes of libraries",
                     format.args = list(big.mark = ",", scientific = FALSE)))
  print(knitr::kable(hto_freq, caption = "Frequency of cells per sample",
                     format.args = list(big.mark = ",", scientific = FALSE)))
  
  
  # Global classification results
  table(hashtag$HTO_classification.global)
  # Group cells based on the max HTO signal
  Idents(hashtag) <- "HTO_maxID"
  
  # cat("#### RidgePlot\n\n")
  # cat("```{r, echo=FALSE, results='asis', warning=FALSE, message=FALSE, dpi=300, fig.width=5, fig.height=6, fig.format='png'}\n")
  
  return(hashtag)
}

```

### GEX library 1 {.tabset}

```{r, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
hashtag <- demultiplex_GEX(GEX_dir=GEX_dir, 
                GEX_name="1_heart_day1", 
                Hashtag_dir=Hashtag_dir, 
                Hashtag_name="Heart_Day1",
                hashtag_labels=c("Day1_Heart_Fed", "Day1_Heart_Fasted", "Day1_Heart_Refed"), 
                hashtag_index=c(1,2,3))
```

#### RidgePlot

```{r, echo=FALSE, results='asis', warning=FALSE, message=FALSE, dpi=300, fig.width=5, fig.height=6, fig.format='png'}
RidgePlot(hashtag, assay = "HTO", features = rownames(hashtag[["HTO"]]), ncol=1) + xlab("Hashtag counts (log10)") 
```


#### HTOHeatmap

```{r, echo=FALSE, results='asis', warning=FALSE, message=FALSE, dpi=300, fig.width=5, fig.height=2, fig.format='png'}
HTOHeatmap(hashtag, assay = "HTO")
```

#### DimPlot

```{r, echo=FALSE, results='asis', warning=FALSE, message=FALSE, dpi=300, fig.width=5, fig.height=3, fig.format='png'}
hashtag <- FindVariableFeatures(hashtag, selection.method = "mean.var.plot")
hashtag <- ScaleData(hashtag, features = VariableFeatures(hashtag))
hashtag <- RunPCA(hashtag)
hashtag <- RunTSNE(hashtag, dims = 1:5, perplexity = 100)
DimPlot(hashtag)
```

#### nCount RNA

```{r, echo=FALSE, results='asis', warning=FALSE, message=FALSE, dpi=300, fig.width=4, fig.height=3, fig.format='png'}
Idents(hashtag) <- "HTO_classification.global"
VlnPlot(hashtag, features = "nCount_RNA", pt.size = 0.1, log = TRUE)
```

