---
title: "GSEA on Hallmark Geneset from MSigDb (directional)"
author:
RMD_AUTHORS
	date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: false
    theme: lumen
bibliography: ../references.bib
---

```{r, echo=FALSE, results="hide", warning=FALSE, message=FALSE}
Sys.setenv(RENV_PATHS_ROOT = "/data/shared_env/renv/")
options(repos = BiocManager::repositories())
Tag_this_analysis <- "GSEA_hallmarks"
pvalueCutoff_GSEA <- 0.01
load("data.RData")

library(dplyr)
library(DT)
library(ggplot2)
library('openxlsx')
library("msigdbr")
library(fgsea)

if (organism == "hsapiens") {
  msig <- msigdbr(species = "Homo sapiens", category = "H")
  pathways = split(x = msig$gene_symbol, f = msig$gs_name)
} else if (organism == "mmusculus") {
  msig <- msigdbr(species = "Mus musculus", category = "H")
  pathways = split(x = msig$gene_symbol, f = msig$gs_name)
}

```

### [Back to front page](Analysis_Report_RNAseq.html)

GSEA[@subramanian2005gene] is a computational method used in genomics to reveal the biological significance of gene expression data. Unlike traditional methods that analyze single genes, GSEA assesses groups of genes, often organized by biological function or pathway. Instead of browsing all available GO terms in the databases, this analysis focuses only the 50 Hallmark Gene Sets from [MSigDB](https://www.gsea-msigdb.org/gsea/msigdb/index.jsp)[@liberzon2011molecular] with the R package fgsea[@korotkevich2016fast].

# Control_vs_Treatment

```{r, echo=FALSE, results='markup', warning=FALSE, message=FALSE}
label="Control_vs_Treatment"

GENE_LIST <- paste0("DGEA_", label, "_genes_stats.csv")
# reading in input from deseq2
df = read.csv(GENE_LIST, header=TRUE)
geneList <- df$log2FoldChange
names(geneList) <- df$gene_name
geneList <- na.omit(geneList)
geneList <- sort(geneList, decreasing = TRUE)

msig_gsea <- fgsea(pathways = pathways,
                   stats = geneList,
                   minSize = 15,   # Set to a positive value for enrichment
                   maxSize = 500,  # Set an appropriate maxSize
                   nperm = 10000,  # Set an appropriate number of permutations
                   gseaParam = 0.5)  # Set an appropriate gseaParam

msig_gsea <- msig_gsea[order(msig_gsea$padj, decreasing = F)]
outdf <- msig_gsea
outdf$leadingEdge <- sapply(outdf$leadingEdge, function(sublist) paste(sublist, collapse = ", "))

write.table(outdf, file = paste0(Tag_this_analysis, "_", label,"_res.csv"))
wb <- createWorkbook()
addWorksheet(wb, "GSEA_stats")
writeData(wb, "GSEA_stats", outdf)
saveWorkbook(wb, file = paste0(Tag_this_analysis, "_", label,"_res.xlsx"), overwrite = TRUE)


DT::datatable(as.data.frame(msig_gsea), extensions = c("FixedColumns"), filter = 'top',
             options = list( autoWidth = TRUE ,
                             dom = 'Blftip',
                             pageLength = 20,
                             searchHighlight = TRUE,
                             scrollX = TRUE,
                             order = list(list(4, 'asc'))
                             ),
             class = c('compact cell-border stripe hover') ,
             rownames = FALSE) %>% formatRound(c("pval","padj","ES", "NES"), 4)
```

Download the full table:

* [`r paste0(Tag_this_analysis, "_", label,"_res.csv")`](`r paste0(Tag_this_analysis, "_", label,"_res.csv")`)
* [`r paste0(Tag_this_analysis, "_", label,"_res.xlsx")`](`r paste0(Tag_this_analysis, "_", label,"_res.xlsx")`)



# GSEA Plot of top gene sets

```{r, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
for (i in 1:50) {
  cat("\n### ",  msig_gsea$pathway[i], "\n")
  p <- plotEnrichment(pathwaysH[[msig_gsea$pathway[i]]], geneList)
  print(p)
  cat("\n")
}
```


# Resources

* [fgsea](https://github.com/ctlab/fgsea)

# References

<div id="refs"></div>

# R session information

```{r, echo=FALSE, results='markup', warning=FALSE, message=TRUE}
sessionInfo()
```