---
title: "Protein-protein interaction analysis by STRING DB"
author:
PROJECT_AUTHORS
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: false
    theme: lumen
    code_folding: hide
---

```{r DNPTGF_vs_TGF-setup, echo=FALSE, results="hide", warning=FALSE, message=FALSE}
source("/etc/rstudio/Rprofile.site")
library(igraph)
library(STRINGdb)
library(viridis)
library(tidyverse)
library(DT)
library(tidygraph)
library(data.table)
library(ggVennDiagram)
library(ggraph)
# library(clusterProfiler)
library(ggplot2)
library(tidyr)
# library(data.table)
library(openxlsx)
library(kableExtra)
options(warn = -1)

Tag_this_analysis <- "stringdb"

CUTOFF_ADJP <- 0.05
CUTOFF_FC <- 0

extract_de_genes <- function(filename, direction = FALSE) {
  degenes <- fread(filename)[, c("gene_name", "log2FoldChange", "padj")]

  if (direction == "up") {
    degenes <- degenes[degenes$log2FoldChange >= CUTOFF_FC]
  } else if (direction == "down") {
    degenes <- degenes[degenes$log2FoldChange <= CUTOFF_FC * -1]
  }
  return(unique(unlist(degenes$gene_name[degenes$padj < CUTOFF_ADJP & (degenes$log2FoldChange >= CUTOFF_FC | degenes$log2FoldChange <= CUTOFF_FC * -1)])))
}

generate_venn <- function(label1, label2, title, direction = FALSE) {
  genes <- list(
    label1 = extract_de_genes(paste0("DGEA_", label1, "_genes_stats.csv"), direction = direction),
    label2 = extract_de_genes(paste0("DGEA_", label2, "_genes_stats.csv"), direction = direction)
  )
  # names(genes) <- c(gsub("_", "", label1),
  #                   gsub("_", "", label2))
  names(genes) <- c(label1, label2)
  # plot(euler(genes), quantities = TRUE)
  p <- gplots::venn(genes, show.plot = F)
  f <- ggVennDiagram(genes) + ggtitle(title) + scale_x_continuous(expand = expansion(mult = .3)) + theme(plot.title = element_text(hjust = 0.5))
  plot(f)
  return(attributes(p)$intersections)
}

get_degenes <- function(sheetname) {
  stat <- read.xlsx(DGEA_export, sheet = sheetname)
  de_genes <- stat$gene_symbol[stat$Significant=="Yes"]
  de_genes <- unique(de_genes)
  return(de_genes)
}

initial_stringDB <- function(input_genes) {
  # Initialize STRINGdb for Homo sapiens (species ID 9606)
  string_db <- STRINGdb$new(
    version = "12.0",        # STRING version (use latest available)
    species = 10116,          # NCBI taxonomy ID for Homo sapiens
    score_threshold = 400,   # Confidence score threshold (range: 0-1000)
    input_directory = ""     # Default; can specify if needed
  )
  # Create a data frame with your gene symbols
  gene_df <- data.frame(gene_symbol = input_genes)
  # Map gene symbols to STRING IDs
  mapped_genes <- string_db$map(gene_df, "gene_symbol", removeUnmappedRows = TRUE )
  # Extract STRING IDs
  string_ids <- mapped_genes$STRING_id
  # Get interactions for the mapped genes
  interactions <- string_db$get_interactions(string_ids)
  interactions <- interactions[duplicated(interactions), ]
  # Rename for 'from' mappings
  mapped_genes_from <- mapped_genes %>%
    dplyr::rename(from = STRING_id, from_gene = gene_symbol)
  # Rename for 'to' mappings
  mapped_genes_to <- mapped_genes %>%
    dplyr::rename(to = STRING_id, to_gene = gene_symbol)
  # Merge 'from' gene symbols
  interactions_named <- interactions %>%
    left_join(mapped_genes_from, by = "from")
  # Merge 'to' gene symbols
  interactions_named <- interactions_named %>%
    left_join(mapped_genes_to, by = "to")
  # Select and arrange the necessary columns
  interactions_named <- interactions_named %>%
    dplyr::select(from_gene, to_gene, combined_score) %>%
    dplyr::arrange(desc(combined_score))
  
  # Create an interactive datatable
  dtable <- datatable(
                interactions_named,
                options = list(
                  pageLength = 10,                 # Number of rows per page
                  autoWidth = TRUE,                # Automatic column width
                  dom = 'Bfrtip',                  # Define the table control elements to appear
                  buttons = c('copy', 'csv', 'excel', 'pdf', 'print')  # Buttons for downloading
                ),
                rownames = FALSE,                  # Do not display row names
                extensions = 'Buttons',            # Enable Buttons extension
              ) %>%
                formatRound(columns = 'combined_score', digits = 0)  # Round combined_score to 2 decimal places
  # Create an igraph object from the interactions
  unique_vertices <- mapped_genes[, c(2,1)] %>% distinct(STRING_id, .keep_all = TRUE)
  network <- graph_from_data_frame(
    d = interactions,
    vertices = unique_vertices,
    directed = FALSE
  )
  
  # Simplify the network (remove multiple edges)
  network <- igraph::simplify(network, remove.multiple = TRUE, remove.loops = FALSE,
                              edge.attr.comb = list(combined_score = "max", "ignore"))
  # Calculate the total combined_score for each node
  node_strength <- strength(network, mode = "all", weights = E(network)$combined_score)
  # Add the node_strength as a vertex attribute
  V(network)$strength <- node_strength
  graph <- as_tbl_graph(network)
  # Plot the network using ggraph
  fig1 <- ggraph(graph, layout = "fr") +  # Fruchterman-Reingold layout
            # Map edge color to combined_score and set a fixed width
            geom_edge_link(aes(color = combined_score), alpha = 0.8) +
            # Set node size based on strength and color as a fixed color
            geom_node_point(aes(size = strength), color = "orange") +
            # Add gene symbol labels with repelling to avoid overlap
            geom_node_text(aes(label = gene_symbol), repel = TRUE, size = 3) +
            # Apply a viridis color scale for edges
            scale_color_viridis(option = "C", name = "Combined Score") +
            # Adjust the size scale for better visibility
            scale_size_continuous(range = c(2, 10), name = "Node Strength") +
            # Remove background and axes for a cleaner look
            theme_void() +
            # Add a title to the plot
            ggtitle("STRING Protein-Protein Interaction Network") +
            # Enhance plot title aesthetics
            theme(
              plot.title = element_text(hjust = 0.5, size = 10, face = "bold")
            )
  # Identify clusters in the network
  clusters <- clusters(network)
  largest_cluster <- which.max(clusters$csize)  # Find the largest cluster
  
  # Filter the network to include only the largest component
  major_cluster_nodes <- V(network)[clusters$membership == largest_cluster]
  major_cluster_network <- induced_subgraph(network, major_cluster_nodes)
  
  # Recalculate node strength for the filtered network
  node_strength_major <- strength(major_cluster_network, mode = "all", weights = E(major_cluster_network)$combined_score)
  V(major_cluster_network)$strength <- node_strength_major
  
  # Convert to a tidygraph format for plotting with ggraph
  graph_major <- as_tbl_graph(major_cluster_network)
  
  # Plot the major cluster network using ggraph
  fig2 <- ggraph(graph_major, layout = "fr") +
            geom_edge_link(aes(color = combined_score), alpha = 0.8) +
            geom_node_point(aes(size = strength), color = "orange") +
            geom_node_text(aes(label = gene_symbol), repel = TRUE, size = 3) +
            scale_color_viridis(option = "C", name = "Combined Score") +
            scale_size_continuous(range = c(2, 10), name = "Node Strength") +
            theme_void() +
            ggtitle("Major Cluster in STRING Protein-Protein Interaction Network") +
            theme(
              plot.title = element_text(hjust = 0.5, size = 10, face = "bold")
            )
  
  node_strength_df <- data.frame(
    STRING_id = V(network)$name,
    gene_symbol = V(network)$gene_symbol,
    node_strength = V(network)$strength,
    stringsAsFactors = FALSE
  ) %>%
    arrange(desc(node_strength))
  # Create an interactive datatable
  nodes <- datatable(
              node_strength_df,
              options = list(
                pageLength = 10,                 # Number of rows per page
                autoWidth = TRUE,                # Automatic column width
                dom = 'Bfrtip',                  # Define the table control elements to appear
                buttons = c('copy', 'csv', 'excel', 'pdf', 'print')  # Buttons for downloading
              ),
              rownames = FALSE,                  # Do not display row names
              extensions = 'Buttons',            # Enable Buttons extension
              caption = htmltools::tags$caption(
                style = 'caption-side: top; text-align: center; font-weight: bold; font-size: 16px;',
                'Table: Gene Nodes Ranked by Strength'
              )
            ) %>%
              formatRound(columns = 'node_strength', digits = 0)  # Round node_strength to 2 decimal places

  return(list("table"=dtable, "fig1"=fig1, "fig2"=fig2, "nodes"=nodes))
}
```

### [Back to front page](../Analysis_Report_RNAseq.html)

This report includes the following steps:

1. Filtering the DE genes Up in TGFvsControl AND Down in TGFDNPvsTGF
2. Further check the fold changes of those genes inÂ DNPvsControl
3. Characterizing this list of genes
    * GO analysis
    * STRING network

## Venn diagram

```{r, echo=TRUE, results="hide", warning=FALSE, message=FALSE}
DGEA_export <- "../DGEA/DGEA_MultiFactor_export.xlsx"
genes <- list(
  WD = get_degenes("WD_Statistics"),
  TAA = get_degenes("TAA_Statistics"),
  WDTAA = get_degenes("WDTAA_Statistics")
)
f <- ggVennDiagram(genes) + scale_x_continuous(expand = expansion(mult = .3)) + 
  scale_fill_gradient(low = "#F4FAFE", high = "#4981BF")
plot(f)
```

# STRING network of DE genes in WD

```{r WD, echo=TRUE, results="asis", warning=FALSE, message=FALSE}
input_genes <- genes[["WD"]]
res <- initial_stringDB(input_genes)
```

### Interactions

```{r, echo=TRUE, results="asis", warning=FALSE, message=FALSE}
res$table
```

### Networks

```{r, echo=TRUE, results="asis", warning=FALSE, message=FALSE}
res$fig1
```

### Majoy cluster

```{r, echo=TRUE, results="asis", warning=FALSE, message=FALSE}
res$fig2
```

### Nodes

```{r, echo=TRUE, results="asis", warning=FALSE, message=FALSE}
res$nodes
```

# R session information

```{r session, echo=FALSE, results='markup', warning=FALSE, message=TRUE}
sessionInfo()
```