---
title: "Quality Control of TITLEDESCRIPTION"
author:
RMD_AUTHORS
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    df_print: paged
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: false
    theme: lumen
bibliography: ../references.bib
---

```{r setup, echo=TRUE, results="hide", warning=FALSE, message=FALSE}
source("/etc/rstudio/Rprofile.site")
# load packages required for analysis
library(knitr)
library(limma)
library(minfi)
library(IlluminaHumanMethylation450kanno.ilmn12.hg19)
library(IlluminaHumanMethylation450kmanifest)
library(RColorBrewer)
library(missMethyl)
library(minfiData)
library(Gviz)
library(DMRcate)
library(ggplot2)
library(stringr)
# options(warn = -1)

Tag_this_analysis <- "1_DNAmArray_QC"
PATH_IDAT <- ""
```

### [Back to front page](../Analysis_Report_DNAmArray.html)

The steps done in this analysis are done according to [A cross-package Bioconductor workflow for analysing methylation array data](https://www.bioconductor.org/packages/release/workflows/vignettes/methylationArrayAnalysis/inst/doc/methylationArrayAnalysis.html)[@maksimovic2016cross].

Maksimovic J, Phipson B, Oshlack A. A cross-package Bioconductor workflow for analysing methylation array data. F1000Res. 2016 Jun 8;5:1281. doi: 10.12688/f1000research.8839.3. PMID: 27347385; PMCID: PMC4916993.

# Loading data

```{r load, echo=TRUE, results="hide", warning=FALSE, message=FALSE}
# get the 450k annotation data
ann450k <- getAnnotation(IlluminaHumanMethylation450kanno.ilmn12.hg19)
# head(ann450k)
# read in the sample sheet for the experiment
targets <- read.metharray.sheet(PATH_IDAT, pattern="samplesheet.csv")
targets <- targets[targets$Sample_Group != "",]
targets <- targets[targets$Note != "dropped",]
# read in the raw data from the IDAT files
rgSet <- read.metharray.exp(base=PATH_IDAT,
                            targets=targets, force=TRUE)
# give the samples descriptive names
targets$ID <- paste(targets$Sample_Group,targets$Sample.Name,sep=".")
sampleNames(rgSet) <- targets$ID
rgSet
```


# Quality control

```{r qc, echo=TRUE, results="hide", warning=FALSE, message=FALSE}
# calculate the detection p-values
detP <- detectionP(rgSet)

qcReport(rgSet, sampNames=targets$ID, sampGroups=targets$Sample_Group, 
         pdf=paste0(Tag_this_analysis,"_Report.pdf"))
```


```{r pfigure, echo=TRUE, results='asis', warning=FALSE, message=FALSE}
# examine mean detection p-values across all samples to identify any failed samples
pal <- brewer.pal(length(unique(targets$Sample_Group)), "Dark2")
# Set up the layout with extra margin space on the right for the legend
par(mar=c(9, 5, 2, 9) + 0.1)
# Create the bar plot
barplot(colMeans(detP), col=pal[factor(targets$Sample_Group)], las=2, 
        cex.names=0.8, ylab="Mean detection p-values")
# Add a horizontal line at p-value 0.05
abline(h=0.05, col="red")
# Add the legend to the right of the plot area
legend("topright", inset=c(-0.37, 0), legend=levels(factor(targets$Sample_Group)), fill=pal, bg="white", xpd=TRUE)
```

```{r filter, echo=TRUE, results='asis', warning=FALSE, message=FALSE}
# remove poor quality samples
keep <- colMeans(detP) < 0.05
rgSet <- rgSet[,keep]
# remove poor quality samples from targets data
targets <- targets[keep,]
```

# Normalisation

```{r normalization, echo=TRUE, results='asis', warning=FALSE, message=FALSE}
# normalize the data; this results in a GenomicRatioSet object
mSetSq <- preprocessQuantile(rgSet)
# create a MethylSet object from the raw data for plotting
mSetRaw <- preprocessRaw(rgSet)
```

# Filtering samples

```{r echo=TRUE, message=FALSE, warning=FALSE, results='asis'}
# Extract current sample IDs from `mSetSq`
# current_samples <- mSetSq$Sample.Name
# Get the sample IDs you want to keep
# keep_groups <- c("ALS", "IBM", "Multiminicores", "NMA", "PM", "Control")
# samples_to_keep <- targets$Sample.Name[targets$Sample_Group %in% keep_groups]
# Subset the `mSetSq` object based on `samples_to_keep`
# mSetSq <- subset(mSetSq, samples_to_keep %in% current_samples)
# mSetSq@colData <- mSetSq@colData[mSetSq@colData$Sample.Name %in% samples_to_keep,]
```

```{r densityplots, echo=TRUE, results='asis', warning=FALSE, message=FALSE, fig.height=6}
# Set up the plotting area to have 2 rows and 1 column (vertical arrangement)
par(mfrow = c(1, 2))

# Plot 1: Raw data
densityPlot(rgSet, sampGroups = targets$Sample_Group, main = "Raw", legend = FALSE)
# Add legend to the right of the first plot
legend("top", legend = levels(factor(targets$Sample_Group)), 
       text.col = brewer.pal(8, "Dark2"), bty = "n")

# Plot 2: Normalized data
densityPlot(getBeta(mSetSq), sampGroups = targets$Sample_Group, main = "Normalized", legend = FALSE)
# Add legend to the right of the second plot
legend("top", legend = levels(factor(targets$Sample_Group)), 
       text.col = brewer.pal(8, "Dark2"), bty = "n")
```

```{r save, echo=TRUE, results='asis', warning=FALSE, message=FALSE}
# Save the objects to a file
save(targets, mSetSq, detP, file = paste0(Tag_this_analysis, ".RData"))
```

# References

<div id="refs"></div>

# R session information

```{r session, echo=TRUE, results='markup', warning=FALSE, message=TRUE}
sessionInfo()
```