---
title: "Processing of TITLEDESCRIPTION"
author:
RMD_AUTHORS
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    df_print: paged
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: false
    theme: lumen
bibliography: ../references.bib
---

```{r setup, echo=TRUE, results="hide", warning=FALSE, message=FALSE}
source("/etc/rstudio/Rprofile.site")
# load packages required for analysis
library(knitr)
library(limma)
library(minfi)
library(IlluminaHumanMethylation450kanno.ilmn12.hg19)
library(IlluminaHumanMethylation450kmanifest)
library(RColorBrewer)
library(missMethyl)
library(minfiData)
library(Gviz)
library(DMRcate)
library(ggplot2)
library(plotly)
library(stringr)
library(methylationArrayAnalysis)
# options(warn = -1)

Tag_this_analysis <- "2_DNAmArray_Processing"

load("1_DNAmArray_QC.RData")
```

### [Back to the main page](../Analysis_Report_DNAmArray.html)

The steps done in this analysis are done according to [A cross-package Bioconductor workflow for analysing methylation array data](https://www.bioconductor.org/packages/release/workflows/vignettes/methylationArrayAnalysis/inst/doc/methylationArrayAnalysis.html)[@maksimovic2016cross].

Maksimovic J, Phipson B, Oshlack A. A cross-package Bioconductor workflow for analysing methylation array data. F1000Res. 2016 Jun 8;5:1281. doi: 10.12688/f1000research.8839.3. PMID: 27347385; PMCID: PMC4916993.

# Before filtering

## PCA (Principal Component Analysis) {.tabset}

PCA seeks to find the directions (principal components) that maximize the variance in the data. It transforms the original variables into a new set of orthogonal variables (principal components) that capture the most variance. The first principal component explains the most variance, and each subsequent component explains the maximum variance possible under the constraint that it is orthogonal to the previous components.

### 2D PCA

```{r, echo=TRUE, results='asis', warning=FALSE, message=FALSE}
# Perform PCA on the M-values
pca_res <- prcomp(t(getM(mSetSq)), scale. = TRUE)
# Prepare data frames for plotly (2D PCA)
df_pca_2d <- data.frame(PC1 = pca_res$x[,1], PC2 = pca_res$x[,2], 
                        Group = factor(targets$Sample_Group),
                        Name = targets$Sample.Name)
# Define colors
pal <- brewer.pal(length(unique(targets$Sample_Group)), "Dark2")
# 2D PCA Plot
pca_plot_2d <- plot_ly(df_pca_2d, x = ~PC1, y = ~PC2, type = 'scatter', mode = 'markers',
              color = ~Group, colors = pal) %>%
  layout(title = 'PCA Plot (2D) by Sample Group',
         xaxis = list(title = paste0('PC1 (', round(100 * summary(pca_res)$importance[2, 1], 2), '% Variance Explained)')),
         yaxis = list(title = paste0('PC2 (', round(100 * summary(pca_res)$importance[2, 2], 2), '% Variance Explained)')))
pca_plot_2d
```

### 3D PCA

```{r, echo=TRUE, results='asis', warning=FALSE, message=FALSE}
# Prepare data frames for plotly (3D PCA)
df_pca_3d <- data.frame(PC1 = pca_res$x[,1], PC2 = pca_res$x[,2], PC3 = pca_res$x[,3], Group = factor(targets$Sample_Group))

# 3D PCA Plot
pca_plot_3d <- plot_ly(df_pca_3d, x = ~PC1, y = ~PC2, z = ~PC3, type = 'scatter3d', mode = 'markers',
              color = ~Group, colors = pal) %>%
  layout(title = 'PCA Plot (3D) by Sample Group',
         scene = list(
           xaxis = list(title = paste0('PC1 (', round(100 * summary(pca_res)$importance[2, 1], 2), '% Variance Explained)')),
           yaxis = list(title = paste0('PC2 (', round(100 * summary(pca_res)$importance[2, 2], 2), '% Variance Explained)')),
           zaxis = list(title = paste0('PC3 (', round(100 * summary(pca_res)$importance[2, 3], 2), '% Variance Explained)'))
         ))
pca_plot_3d
```

## MDS (Multidimensional Scaling)) {.tabset}

MDS focuses on preserving the pairwise distances or dissimilarities between data points. It tries to position the points in a lower-dimensional space such that the distances between points in the plot are as close as possible to the original distances in the high-dimensional space.

### 2D MDS

```{r, echo=TRUE, results='asis', warning=FALSE, message=FALSE}
# Generate MDS coordinates using cmdscale
mds_res <- cmdscale(dist(t(getM(mSetSq))), k = 2) # k = 2 for 2D MDS
# Prepare data frames for plotly (2D MDS)
df_mds_2d <- data.frame(Dim1 = mds_res[,1], Dim2 = mds_res[,2], Group = factor(targets$Sample_Group))
# Define colors
pal <- brewer.pal(length(unique(targets$Sample_Group)), "Dark2")
# 2D MDS Plot
mds_plot_2d <- plot_ly(df_mds_2d, x = ~Dim1, y = ~Dim2, type = 'scatter', mode = 'markers',
              color = ~Group, colors = pal) %>%
  layout(title = 'MDS Plot (2D) by Sample Group',
         xaxis = list(title = 'Dimension 1'),
         yaxis = list(title = 'Dimension 2'))
mds_plot_2d
```

### 3D MDS

```{r, echo=TRUE, results='asis', warning=FALSE, message=FALSE}
# Generate MDS coordinates using cmdscale
mds_res_3d <- cmdscale(dist(t(getM(mSetSq))), k = 3) # k = 3 for 3D MDS

# Prepare data frames for plotly (3D MDS)
df_mds_3d <- data.frame(Dim1 = mds_res_3d[,1], Dim2 = mds_res_3d[,2], Dim3 = mds_res_3d[,3], Group = factor(targets$Sample_Group))

# 3D MDS Plot
mds_plot_3d <- plot_ly(df_mds_3d, x = ~Dim1, y = ~Dim2, z = ~Dim3, type = 'scatter3d', mode = 'markers',
              color = ~Group, colors = pal) %>%
  layout(title = 'MDS Plot (3D) by Sample Group',
         scene = list(
           xaxis = list(title = 'Dimension 1'),
           yaxis = list(title = 'Dimension 2'),
           zaxis = list(title = 'Dimension 3')
         ))
mds_plot_3d

```

## Calculating M-values and beta values

```{r, echo=TRUE, results='asis', warning=FALSE, message=FALSE}
# calculate M-values for statistical analysis
mVals <- getM(mSetSq)
bVals <- getBeta(mSetSq)
```

```{r, echo=TRUE, results='asis', warning=FALSE, message=FALSE}
# Increase the right margin to make space for the legend
par(mar=c(5, 4, 4, 8) + 0.1)
# Plot the density plot
densityPlot(bVals, sampGroups=targets$Sample_Group, main="Beta values", 
            legend=FALSE, xlab="Beta values")
# Place the legend outside the plot area
legend("topright", inset=c(-0.3,0), legend = levels(factor(targets$Sample_Group)), 
       text.col=brewer.pal(8,"Dark2"), xpd=TRUE, bty = "n")
```


```{r, echo=TRUE, results='asis', warning=FALSE, message=FALSE}
# Increase the right margin to make space for the legend
par(mar=c(5, 4, 4, 8) + 0.1)
# Plot the density plot
densityPlot(mVals, sampGroups=targets$Sample_Group, main="M values", 
            legend=FALSE, xlab="M values")
# Place the legend outside the plot area
legend("topright", inset=c(-0.3,0), legend = levels(factor(targets$Sample_Group)), 
       text.col=brewer.pal(8,"Dark2"), xpd=TRUE, bty = "n")
```

# Filtering probes

## Detection p-value

Poor performing probes are generally filtered out prior to differential methylation analysis. As the signal from these probes is unreliable, by removing them we perform fewer statistical tests and thus incur a reduced multiple testing penalty. We filter out probes that have failed in one or more samples based on detection p-value.

```{r, echo=TRUE, results='asis', warning=FALSE, message=FALSE}
# ensure probes are in the same order in the mSetSq and detP objects
detP <- detP[match(featureNames(mSetSq),rownames(detP)),]
# remove any probes that have failed in one or more samples
keep <- rowSums(detP < 0.01) == ncol(mSetSq) 
mSetSq <- mSetSq[keep,]
```

|Probes | Number |
|------------- | ------------:|
|Total before filtering | `r length(keep)`|
|To be removed  | `r sum(!keep)`|
|To be kept  | `r sum(keep)`|

## Probes on the X and Y chromosomes

Depending on the nature of your samples and your biological question you may also choose to filter out the probes from the X and Y chromosomes or probes that are known to have common SNPs at the CpG site.

```{r, echo=TRUE, results='asis', warning=FALSE, message=FALSE}
# get the 450k annotation data
ann450k <- getAnnotation(IlluminaHumanMethylation450kanno.ilmn12.hg19)
# if your data includes males and females, remove probes on the sex chromosomes
keep <- !(featureNames(mSetSq) %in% ann450k$Name[ann450k$chr %in% c("chrX","chrY")])
mSetSq <- mSetSq[keep,]
```

|Probes | Number |
|------------- | ------------:|
|Total before filtering | `r length(keep)`|
|To be removed  | `r sum(!keep)`|
|To be kept  | `r sum(keep)`|

## SNPs

```{r, echo=TRUE, results='asis', warning=FALSE, message=FALSE}
# remove probes with SNPs at CpG site
mSetSq <- dropLociWithSnps(mSetSq)
```

|Probes | Number |
|------------- | ------------:|
|Total before filtering | `r length(keep)`|
|Total after filtering | `r dim(mSetSq)[1]`|

## Cross-reactive CpGs

We will also filter out probes that have shown to be cross-reactive, that is, probes that have been demonstrated to map to multiple places in the genome. This list was originally published by Chen et al. (2013) and can be obtained from [the authors’ website](http://www.sickkids.ca/MS-Office-Files/Research/Weksberg%20Lab/48639-non-specific-probes-Illumina450k.xlsx).

```{r, echo=TRUE, results='asis', warning=FALSE, message=FALSE}
# exclude cross reactive probes
dataDirectory <- system.file("extdata", package = "methylationArrayAnalysis")
xReactiveProbes <- read.csv(file=paste(dataDirectory,
                                       "48639-non-specific-probes-Illumina450k.csv",
                                       sep="/"), stringsAsFactors=FALSE)
keep <- !(featureNames(mSetSq) %in% xReactiveProbes$TargetID)
mSetSq <- mSetSq[keep,]
```

| Probes | Number |
| ------------- | ------------:|
| Total before filtering | `r length(keep)`|
| To be removed  | `r sum(!keep)`|
| To be kept  | `r sum(keep)`|

# After filtering

### Filtered t-SNE

```{r, echo=TRUE, results='asis', warning=FALSE, message=FALSE}
# Load necessary library
library(Rtsne)

# Perform t-SNE on the transposed M-values (same input as PCA)
set.seed(42)  # Set a seed for reproducibility
tsne_res <- Rtsne(t(getM(mSetSq)), perplexity = 15, theta = 0.5, dims = 2)

# Prepare data frames for plotly (2D t-SNE)
df_tsne_2d <- data.frame(
  TSNE1 = tsne_res$Y[,1],
  TSNE2 = tsne_res$Y[,2],
  Group = factor(targets$Sample_Group)
)

# Define colors (same palette as used in PCA plot)
pal <- brewer.pal(length(unique(targets$Sample_Group)), "Set1")

# 2D t-SNE Plot
tsne_plot_2d <- plot_ly(df_tsne_2d, x = ~TSNE1, y = ~TSNE2, type = 'scatter', mode = 'markers',
              color = ~Group, colors = pal) %>%
  layout(title = 't-SNE Plot (2D) by Sample Group',
         xaxis = list(title = 't-SNE1'),
         yaxis = list(title = 't-SNE2'))

# Show the plot
tsne_plot_2d

```

## Filtered PCA (Principal Component Analysis) {.tabset}

PCA seeks to find the directions (principal components) that maximize the variance in the data. It transforms the original variables into a new set of orthogonal variables (principal components) that capture the most variance. The first principal component explains the most variance, and each subsequent component explains the maximum variance possible under the constraint that it is orthogonal to the previous components.

### Filtered 2D PCA

```{r, echo=TRUE, results='asis', warning=FALSE, message=FALSE}
# Perform PCA on the M-values
pca_res <- prcomp(t(getM(mSetSq)), scale. = TRUE)
# Prepare data frames for plotly (2D PCA)
df_pca_2d <- data.frame(PC1 = pca_res$x[,1], PC2 = pca_res$x[,2], Group = factor(targets$Sample_Group))
# Define colors
pal <- brewer.pal(length(unique(targets$Sample_Group)), "Dark2")
# 2D PCA Plot
pca_plot_2d <- plot_ly(df_pca_2d, x = ~PC1, y = ~PC2, type = 'scatter', mode = 'markers',
              color = ~Group, colors = pal) %>%
  layout(title = 'PCA Plot (2D) by Sample Group',
         xaxis = list(title = paste0('PC1 (', round(100 * summary(pca_res)$importance[2, 1], 2), '% Variance Explained)')),
         yaxis = list(title = paste0('PC2 (', round(100 * summary(pca_res)$importance[2, 2], 2), '% Variance Explained)')))
pca_plot_2d
```

### Filtered 3D PCA

```{r, echo=TRUE, results='asis', warning=FALSE, message=FALSE}
# Prepare data frames for plotly (3D PCA)
df_pca_3d <- data.frame(PC1 = pca_res$x[,1], PC2 = pca_res$x[,2], PC3 = pca_res$x[,3], Group = factor(targets$Sample_Group))
# 3D PCA Plot
pca_plot_3d <- plot_ly(df_pca_3d, x = ~PC1, y = ~PC2, z = ~PC3, type = 'scatter3d', mode = 'markers',
              color = ~Group, colors = pal) %>%
  layout(title = 'PCA Plot (3D) by Sample Group',
         scene = list(
           xaxis = list(title = paste0('PC1 (', round(100 * summary(pca_res)$importance[2, 1], 2), '% Variance Explained)')),
           yaxis = list(title = paste0('PC2 (', round(100 * summary(pca_res)$importance[2, 2], 2), '% Variance Explained)')),
           zaxis = list(title = paste0('PC3 (', round(100 * summary(pca_res)$importance[2, 3], 2), '% Variance Explained)'))
         ))
pca_plot_3d
```

## Filtered MDS (Multidimensional Scaling) {.tabset}

MDS focuses on preserving the pairwise distances or dissimilarities between data points. It tries to position the points in a lower-dimensional space such that the distances between points in the plot are as close as possible to the original distances in the high-dimensional space.

### Filtered 2D MDS

```{r, echo=TRUE, results='asis', warning=FALSE, message=FALSE}
# Generate MDS coordinates using cmdscale
mds_res <- cmdscale(dist(t(getM(mSetSq))), k = 2) # k = 2 for 2D MDS
# Prepare data frames for plotly (2D MDS)
df_mds_2d <- data.frame(Dim1 = mds_res[,1], Dim2 = mds_res[,2], Group = factor(targets$Sample_Group))
# Define colors
pal <- brewer.pal(length(unique(targets$Sample_Group)), "Dark2")
# 2D MDS Plot
mds_plot_2d <- plot_ly(df_mds_2d, x = ~Dim1, y = ~Dim2, type = 'scatter', mode = 'markers',
              color = ~Group, colors = pal) %>%
  layout(title = 'MDS Plot (2D) by Sample Group',
         xaxis = list(title = 'Dimension 1'),
         yaxis = list(title = 'Dimension 2'))
mds_plot_2d
```

### Filtered 3D MDS

```{r, echo=TRUE, results='asis', warning=FALSE, message=FALSE}
# Generate MDS coordinates using cmdscale
mds_res_3d <- cmdscale(dist(t(getM(mSetSq))), k = 3) # k = 3 for 3D MDS
# Prepare data frames for plotly (3D MDS)
df_mds_3d <- data.frame(Dim1 = mds_res_3d[,1], Dim2 = mds_res_3d[,2], Dim3 = mds_res_3d[,3], Group = factor(targets$Sample_Group))
# 3D MDS Plot
mds_plot_3d <- plot_ly(df_mds_3d, x = ~Dim1, y = ~Dim2, z = ~Dim3, type = 'scatter3d', mode = 'markers',
              color = ~Group, colors = pal) %>%
  layout(title = 'MDS Plot (3D) by Sample Group',
         scene = list(
           xaxis = list(title = 'Dimension 1'),
           yaxis = list(title = 'Dimension 2'),
           zaxis = list(title = 'Dimension 3')
         ))
mds_plot_3d
```

## Calculating M-values and beta values

```{r, echo=TRUE, results='asis', warning=FALSE, message=FALSE}
# calculate M-values for statistical analysis
mVals <- getM(mSetSq)
bVals <- getBeta(mSetSq)
```

```{r, echo=TRUE, results='asis', warning=FALSE, message=FALSE}
# Increase the right margin to make space for the legend
par(mar=c(5, 4, 4, 8) + 0.1)
# Plot the density plot
densityPlot(bVals, sampGroups=targets$Sample_Group, main="Beta values", 
            legend=FALSE, xlab="Beta values")
# Place the legend outside the plot area
legend("topright", inset=c(-0.3,0), legend = levels(factor(targets$Sample_Group)), 
       text.col=brewer.pal(8,"Dark2"), xpd=TRUE, bty = "n")
```


```{r, echo=TRUE, results='asis', warning=FALSE, message=FALSE}
# Increase the right margin to make space for the legend
par(mar=c(5, 4, 4, 8) + 0.1)
# Plot the density plot
densityPlot(mVals, sampGroups=targets$Sample_Group, main="M values", 
            legend=FALSE, xlab="M values")
# Place the legend outside the plot area
legend("topright", inset=c(-0.3,0), legend = levels(factor(targets$Sample_Group)), 
       text.col=brewer.pal(8,"Dark2"), xpd=TRUE, bty = "n")
```

# Save objects

```{r save, echo=TRUE, results='asis', warning=FALSE, message=FALSE}
write.csv(bVals, file = "processed_bVals.csv")
write.csv(mVals, file = "processed_mVals.csv")
# Save the objects to a file
save(targets, bVals, mVals, file = paste0(Tag_this_analysis, ".RData"))
```

# References

<div id="refs"></div>

# R session information

```{r session, echo=TRUE, results='markup', warning=FALSE, message=TRUE}
sessionInfo()
```