---
title: "Differential Analysis of TITLEDESCRIPTION"
author:
RMD_AUTHORS
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    df_print: paged
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: false
    theme: lumen
bibliography: ../references.bib
---

```{r setup, echo=TRUE, results="hide", warning=FALSE, message=FALSE}
source("/etc/rstudio/Rprofile.site")
# load packages required for analysis
library(knitr)
library(limma)
library(RColorBrewer)
library(pheatmap)
library(DT)
library(openxlsx)
# options(warn = -1)

Tag_this_analysis <- "3_DNAmArray_Differential_Analysis"

load("2_DNAmArray_Processing.RData")

# Define the threshold for significance (e.g., adjusted p-value < 0.05)
alpha <- 0.001
```

### [Back to the main page](../Analysis_Report_DNAmArray.html)

The steps done in this analysis are done according to [A cross-package Bioconductor workflow for analysing methylation array data](https://www.bioconductor.org/packages/release/workflows/vignettes/methylationArrayAnalysis/inst/doc/methylationArrayAnalysis.html)[@maksimovic2016cross].

> Maksimovic J, Phipson B, Oshlack A. A cross-package Bioconductor workflow for analysing methylation array data. F1000Res. 2016 Jun 8;5:1281. doi: 10.12688/f1000research.8839.3. PMID: 27347385; PMCID: PMC4916993.

The threshold applied on the adjusted p-value is `r alpha`

# Sample Group Counts

```{r, echo=TRUE, results='asis', warning=FALSE, message=FALSE}
# Create a contingency table
group_counts <- table(targets$Sample_Group)
# Convert the table to a data frame
group_counts_df <- as.data.frame(group_counts)
# Rename the columns to "Group" and "Number"
colnames(group_counts_df) <- c("Group", "Number")
datatable(group_counts_df, 
          options = list(pageLength = nrow(group_counts_df)),
          rownames = FALSE,  # Removes row names if not needed
          caption = 'Sample Group Counts')
```

# Differenial Analysis

We performed the differential analysis on each sample group comparing to the control samples.

```{r, echo=TRUE, results='asis', warning=FALSE, message=FALSE}
# This is the factor of interest
groups <- factor(targets$Sample_Group)
# This is the individual effect that we need to account for
# individual <- factor(targets$Sample_Source) 
# Use the above to create a design matrix
design <- model.matrix(~0 + groups, data=targets)
colnames(design) <- c(levels(groups))

# Create a contrast matrix for "one vs control" comparisons
contMatrix <- makeContrasts(
  ALS_vs_Control = ALS - Control,
  IBM_vs_Control = IBM - Control,
  levels = design
)

```


```{r, echo=TRUE, results='asis', warning=FALSE, message=FALSE}
# Fit the linear model
fit <- lmFit(mVals, design)
# Apply the contrast matrix
fit2 <- contrasts.fit(fit, contMatrix)
# Apply empirical Bayes smoothing
fit2 <- eBayes(fit2)
```

```{r, echo=TRUE, results='asis', warning=FALSE, message=FALSE}
# Initialize a vector to store the results
comparison_names <- colnames(contMatrix)
significant_cpgs <- numeric(length(comparison_names))
# Loop through each comparison and count the number of significant CpGs
for (i in seq_along(comparison_names)) {
  top_table <- topTable(fit2, coef=comparison_names[i], number=Inf, adjust.method="BH")
  significant_cpgs[i] <- sum(top_table$adj.P.Val < alpha)
}
# Create a data frame to summarize the results
summary_table <- data.frame(
  Comparison = comparison_names,
  Significant_CpGs = significant_cpgs
)
# Print the summary table
datatable(summary_table, 
          options = list(pageLength = nrow(summary_table)),
          rownames = FALSE,  # Removes row names if not needed
          caption = 'Number of Sig. Probes')
```

## Heatmap of top 500 Sig. Probes in all comparisons

```{r, echo=TRUE, results='asis', warning=FALSE, message=FALSE, fig.height=10, fig.width=12}
Num_top_CpGs <- 500
# Initialize an empty list to store top CpGs for each comparison
top_cpgs_list <- list()

# Loop through each comparison and extract the top CpGs
for (i in colnames(contMatrix)) {
  top_cpgs <- topTable(fit2, coef=i, number=Num_top_CpGs, adjust.method="BH")
  top_cpgs_list[[i]] <- rownames(top_cpgs)
}

# Combine all top CpGs into one vector (some CpGs might overlap across comparisons)
all_top_cpgs <- unique(unlist(top_cpgs_list))

# Subset your original data matrix (e.g., mVals or beta values) to include only the top CpGs
heatmap_data <- mVals[all_top_cpgs, ]

# Create the column annotation data frame
column_annotation <- data.frame(
  Group = factor(targets$Sample_Group)
)
rownames(column_annotation) <- colnames(mVals)  # Assuming colnames(mVals) are the sample IDs

# Annotation colors
group_colors <- brewer.pal(12, "Paired")[1:length(unique(column_annotation$Group))]
# Create a named vector to map groups to colors
group_color_mapping <- setNames(group_colors, levels(column_annotation$Group))

# Combine the annotation colors (only for column annotations now)
annotation_colors <- list(
  Group = group_color_mapping  # For the column annotations (sample groups)
)

# Generate heatmap
pheatmap(heatmap_data,
         scale = "row",                       # Standardize each CpG (row) across samples
         clustering_distance_rows = "euclidean",
         clustering_distance_cols = "euclidean",
         clustering_method = "complete",
         show_rownames = FALSE,               # Optionally hide CpG names if too many
         show_colnames = TRUE,                # Show sample names
         main = "Heatmap of Top 100 CpGs in Each Comparison",
         annotation_col = column_annotation,  # Add column annotation
         annotation_colors = annotation_colors # Apply custom colors
)

```

# Save data into an excel file

Please download the full statistics here: [`r paste0(Tag_this_analysis, "_statistics.xlsx")`](`r paste0(Tag_this_analysis, "_statistics.xlsx")`)

```{r, echo=TRUE, results='asis', warning=FALSE, message=FALSE}
# Create a new workbook
wb <- createWorkbook()
# Loop through each comparison, save top_table to a sheet
for (i in seq_along(comparison_names)) {
  top_table <- topTable(fit2, coef=comparison_names[i], number=Inf, adjust.method="BH")
  # Add a new sheet with the name of the comparison
  addWorksheet(wb, sheetName = comparison_names[i])
  # Write the top_table data frame to the worksheet
  writeData(wb, sheet = comparison_names[i], top_table)
}
# Save the workbook to a file
saveWorkbook(wb, file = paste0(Tag_this_analysis, "_statistics.xlsx"), overwrite = TRUE)
```

# Save objects

```{r save, echo=TRUE, results='asis', warning=FALSE, message=FALSE}
sel_mValue <- mVals[all_top_cpgs,]
sel_bValue <- bVals[all_top_cpgs,]
# Save the objects to a file
save(comparison_names, fit2, sel_mValue, sel_bValue, file = paste0(Tag_this_analysis, ".RData"))
```

# References

<div id="refs"></div>

# R session information

```{r session, echo=TRUE, results='markup', warning=FALSE, message=TRUE}
sessionInfo()
```