---
title: "DGEA Geneset Comparisons"
author:
RMD_AUTHORS
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: false
    theme: lumen
bibliography: ../references.bib
---

```{r, echo=FALSE, results="hide", warning=FALSE, message=FALSE}
source("/etc/rstudio/Rprofile.site")
options(repos = BiocManager::repositories())

library(data.table)
library(ggplot2)
library(ggpubr)
library(ggVennDiagram)
library(UpSetR)
library(DT)
library(knitr)

# source("report_functions.R")
# load("data.RData")

CUTOFF_ADJP <- 0.05
CUTOFF_FC <- 1
COUNTER <- 0 # Counter to avoid the colliding of chunk names


Tag_this_analysis <- "DGEA_Geneset_Comparisons"

extract_de_genes <- function(filename, direction = FALSE) {
  degenes <- fread(filename)[, c("gene_name", "log2FoldChange", "padj")]

  if (direction == "up") {
    degenes <- degenes[degenes$log2FoldChange >= CUTOFF_FC]
  } else if (direction == "down") {
    degenes <- degenes[degenes$log2FoldChange <= CUTOFF_FC * -1]
  }
  return(unique(unlist(degenes$gene_name[degenes$padj < CUTOFF_ADJP & (degenes$log2FoldChange >= CUTOFF_FC | degenes$log2FoldChange <= CUTOFF_FC * -1)])))
}

generate_venn <- function(label1, label2, title, direction = FALSE) {
  genes <- list(
    label1 = extract_de_genes(paste0("DGEA_", label1, "_genes_stats.csv"), direction = direction),
    label2 = extract_de_genes(paste0("DGEA_", label2, "_genes_stats.csv"), direction = direction)
  )
  # names(genes) <- c(gsub("_", "", label1),
  #                   gsub("_", "", label2))
  names(genes) <- c(label1, label2)
  # plot(euler(genes), quantities = TRUE)
  p <- gplots::venn(genes, show.plot = F)
  f <- ggVennDiagram(genes) + ggtitle(title) + scale_x_continuous(expand = expansion(mult = .3)) + theme(plot.title = element_text(hjust = 0.5))
  plot(f)
  return(attributes(p)$intersections)
}

generate_venn3 <- function(label1, label2, label3, title, direction = FALSE) {
  genes <- list(
    label1 = extract_de_genes(paste0("DGEA_", label1, "_genes_stats.csv"), direction = direction),
    label2 = extract_de_genes(paste0("DGEA_", label2, "_genes_stats.csv"), direction = direction),
    label3 = extract_de_genes(paste0("DGEA_", label3, "_genes_stats.csv"), direction = direction)
  )
  # names(genes) <- c(gsub("_", "", label1),
  #                   gsub("_", "", label2))
  names(genes) <- c(label1, label2, label3)
  # plot(euler(genes), quantities = TRUE)
  p <- gplots::venn(genes, show.plot = F)
  f <- ggVennDiagram(genes) + ggtitle(title) + scale_x_continuous(expand = expansion(mult = .3)) + theme(plot.title = element_text(hjust = 0.5))
  plot(f)
  return(attributes(p)$intersections)
}


venn_2_excel <- function(name, venn_components) {
  require(openxlsx)
  names_list <- venn_components
  sheet_names <- c("A", "B", "C", "D", "E", "G", "H")
  wb <- openxlsx::createWorkbook()
  for (i in seq_len(length(venn_components))) {
    addWorksheet(wb = wb, sheetName = sheet_names[i])
    writeData(wb = wb, sheet = sheet_names[i], x = names(names_list[i]), startRow = 1, colNames = T, rowNames = T)
    writeData(wb = wb, sheet = sheet_names[i], x = names_list[[i]], startRow = 2, colNames = T, rowNames = T, name = names(names_list[i]))
  }
  saveWorkbook(wb, paste0(name, ".xlsx"), overwrite = TRUE)
}


expand_tables = function(venn_components){

    out = NULL
    for (i in seq_len(length(venn_components))) {
      name = names(venn_components[i])
      knit_expanded <- paste0("\n#### ",i,
                              "\n```{r '",COUNTER,"',  results='asis', echo=FALSE}\n
                              datatable(as.data.frame(venn_components[[",i,"]]), colnames =c('",name,"'))\n
                              \n
                              ```")
      out = c(out, knit_expanded)
      COUNTER = COUNTER + 1
    }
    return(out)
}


```



### [Back to front page](../Analysis_Report_RNAseq.html)

## All DE genes

```{r, echo=FALSE, results="asis", warning=FALSE, message=FALSE, dpi=600, fig.width=6, fig.height=4}
venn_components <- generate_venn3("CSK_vs_SF", "MyoSF_vs_SF", "MyoSF_vs_CSK")
venn_2_excel("save_name", venn_components)
```


<!-- Use if needed to add embedded tables for the venn diagram components -->
### Tables {.tabset}
```{r, echo=FALSE}
venn_tables_rmd = expand_tables(venn_components)
```
<!-- Workaround to expand and knit the resulting string -->
`r paste(knit(text = venn_tables_rmd), collapse = '\n')`

## Up genes

```{r, echo=FALSE, results="asis", warning=FALSE, message=FALSE, dpi=600, fig.width=6, fig.height=4}
venn_components <- generate_venn3("CSK_vs_SF", "MyoSF_vs_SF", "MyoSF_vs_CSK", direction = "up")
venn_2_excel("save_name", venn_components)
```

<!-- Use if needed to add embedded tables for the venn diagram components -->
### Tables {.tabset}
```{r, echo=FALSE}
venn_tables_rmd = expand_tables(venn_components)
```
<!-- Workaround to expand and knit the resulting string -->
`r paste(knit(text = venn_tables_rmd), collapse = '\n')`


## Down genes

```{r, echo=FALSE, results="asis", warning=FALSE, message=FALSE, dpi=600, fig.width=6, fig.height=4}
venn_components = generate_venn3("CSK_vs_SF", "MyoSF_vs_SF", "MyoSF_vs_CSK", direction = "down")
venn_2_excel("save_name", venn_components)
```


<!-- Use if needed to add embedded tables for the venn diagram components -->
### Tables {.tabset}
```{r, echo=FALSE}
venn_tables_rmd = expand_tables(venn_components)
```
<!-- Workaround to expand and knit the resulting string -->
`r paste(knit(text = venn_tables_rmd), collapse = '\n')`


# References

<div id="refs"></div>


# R session information

```{r, echo=FALSE, results='markup', warning=FALSE, message=TRUE}
sessionInfo()
```
