---
title: "DGEA Correlation Analysis"
author:
RMD_AUTHORS
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: false
    theme: lumen
bibliography: ../references.bib
---

```{r, echo=FALSE, results="hide", warning=FALSE, message=FALSE}
Sys.setenv(PATH = paste(Sys.getenv("PATH"), "/opt/miniconda3/envs/R4.3.2/bin", sep = ":"))
Sys.setenv(RENV_PATHS_ROOT = "/data/shared_env/renv/")
options(repos = BiocManager::repositories())
library(data.table)
require(dplyr)
require(DT)
library(ggplot2)
library(ggpubr)
library(pheatmap)
library(tidyr)
options(warn = -1)

source("report_functions.R")

load("data.RData")

Tag_this_analysis <- "DGEA_Correlation_analysis"

corr_figure <- function(title, xlab, ylab) {
  xdata <- fread(paste0("DGEA_", xlab, "_genes_stats.csv"))[, c("gene_id","gene_name","log2FoldChange")]
  ydata <- fread(paste0("DGEA_", ylab, "_genes_stats.csv"))[, c("gene_id","gene_name","log2FoldChange")]
  
  merged_table <- merge(xdata, ydata, 
                        by=c("gene_id","gene_name"), suffixes = c(".xlab", ".ylab"))
  yrange <- abs(max(ydata$log2FoldChange)-min(ydata$log2FoldChange))
  statcor_y <- max(ydata$log2FoldChange)
  statregline_y <- max(ydata$log2FoldChange)-0.05 * yrange
  fig <- ggplot(merged_table, aes(log2FoldChange.xlab, log2FoldChange.ylab)) +
         geom_point(size=1, color="royalblue", alpha=0.2) + theme_bw() +
         ggtitle(title) + theme(plot.title = element_text(hjust = 0.5)) +
         geom_smooth(method = "lm", color="gray", alpha=0.1) +
         xlab(xlab) + ylab(ylab) +
         stat_cor(label.y = statcor_y)+ #this means at 35th unit in the y axis, the r squared and p value will be shown
         stat_regline_equation(label.y = statregline_y) #this means at 30th unit regresion line equation will be shown
  
  fig
}
```

### [Back to front page](../Analysis_Report_RNAseq.html)

If there is a need to modify the figures to fit your particular need (style or color), additional service for customized visualization is needed. The main focus of this report is to present the analysis result to the clients. Any requests beyond this purpose is regarded another service. 

# Correlation of the effect of IL10 and LPS {.tabset}

## WT

```{r, echo=FALSE, results="asis", warning=FALSE, message=FALSE, dpi=600, fig.width=4, fig.height=4}
corr_figure("Correlation of IL10 and LPS", "IL10_vs_Control_in_WT", "LPS_vs_Control_in_WT")
```

## KO

```{r, echo=FALSE, results="asis", warning=FALSE, message=FALSE, dpi=600, fig.width=4, fig.height=4}
corr_figure("Correlation of IL10 and LPS", "IL10_vs_Control_in_KO", "LPS_vs_Control_in_KO")
```

# References

<div id="refs"></div>


# R session information

```{r, echo=FALSE, results='markup', warning=FALSE, message=TRUE}
sessionInfo()
```