---
params:
  authors:
  paired:
  filetag:
  samplesheet:
title: "Differential Gene Expression Analysis for `r gsub('_', ' ', params$filetag)`"
author:
  - "`r params$authors`"
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: false
    theme: lumen
    code_download: true
    code_folding: hide
bibliography: ../references.bib
---

```{r setup, echo=TRUE, warning=FALSE, message=FALSE}
source("/etc/rstudio/Rprofile.site")
# Load libraries
library(DESeq2)
library(dplyr)
library(ggplot2)
library(plotly)
library(viridis) 
library(DT)
library(tximport)
library(data.table)
library(openxlsx)
library(ggrepel)
library(tibble)
library(RColorBrewer)
library(pheatmap)

load("DGEA_params.RData")

description_parts <- strsplit(gsub('_', ' ', params$filetag), " ")[[1]]
samplesheet$group <- factor(samplesheet$group, levels = c(description_parts[3], description_parts[1]))
samplesheet <- samplesheet[complete.cases(samplesheet$group),]
```

### [Back to front page](../Analysis_Report_RNAseq.html)

# DESeq2

Differential gene expression analysis is done with DESEQ2 package in R [@love2014moderated] and follows the instruction for Salmon quantification ([Analyzing RNA-seq data with DESeq2](https://bioconductor.org/packages/release/bioc/vignettes/DESeq2/inst/doc/DESeq2.html)).

```{r DESeq2, echo=TRUE, warning=FALSE, message=FALSE}
# Load transcript-to-gene mapping
tx2gene <- fread(tx2gene_file, col.names = c("transcript_id", "gene_id", "gene_name"))
# Run DESeq2 Analysis
files <- file.path(salmon_dir, samplesheet$sample, "quant.sf")
names(files) <- samplesheet$sample
txi <- tximport(files, type = "salmon", tx2gene = tx2gene[, c(1, 2)],
                countsFromAbundance = counts_from_abundance)

# Set up DESeq2 design formula
if (length_correction) {
  design_formula <- if (params$paired) ~ batch + group else ~ group
  dds <- DESeqDataSetFromTximport(txi, colData = samplesheet, design = design_formula)
} else {
  dds <- DESeqDataSetFromMatrix(round(txi$counts), samplesheet, ~ group)
}

# Perform DESeq2 differential expression analysis
dds <- DESeq(dds)
res <- results(dds, alpha = cutoff_adj_p)
res <- as.data.frame(res)
res <- res[complete.cases(res), ]
res$gene_symbol <- tx2gene$gene_name[match(rownames(res), tx2gene$gene_id)]
res$gene_id <- rownames(res)
res$Significant <- ifelse(res$padj < cutoff_adj_p, "Yes", "No")
res <- res[, c("gene_id", "gene_symbol", "baseMean", "log2FoldChange", "lfcSE", "stat", "pvalue", "padj", "Significant")]
# Get normalized read counts
normalized_counts <- counts(dds, normalized=TRUE)
out_norm <- merge(tx2gene[,-1], rownames_to_column(as.data.frame(normalized_counts), var = "gene_id"),
                  by.x="gene_id", by.y="gene_id")
out_norm <- distinct(out_norm)
# Calculate PCA on normalized counts
pca <- prcomp(t(normalized_counts))  # Transpose to have samples as rows
```

# Interactive figures {.tabset}

## 3D PCA

```{r plotly-3d-pca, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
# Calculate explained variance for PC1, PC2, and PC3
explained_variance <- round(100 * (pca$sdev^2 / sum(pca$sdev^2)), 2)
explained_variance_text <- paste0("PC", 1:3, " (", explained_variance[1:3], "%)")

# Extract scores for the first three principal components
pca_data <- as.data.frame(pca$x[, 1:3])  # Select only the first 3 PCs
pca_data$sample <- rownames(pca_data)  # Add sample names as a column
pca_data$group <- samplesheet$group    # Add group information from samplesheet

# Create 3D PCA plot with plotly
fig <- plot_ly(
  data = pca_data,
  x = ~PC1, y = ~PC2, z = ~PC3,
  color = ~group,
  colors = "Dark2",  # Use a color palette suitable for publication
  text = ~paste("Sample:", sample),  # Add sample names as hover text
  type = "scatter3d",
  mode = "markers",
  marker = list(size = 9)  # Set marker size for visibility
) %>%
  layout(
    title = "3D PCA Plot of Samples",
    scene = list(
      xaxis = list(title = explained_variance_text[1]),
      yaxis = list(title = explained_variance_text[2]),
      zaxis = list(title = explained_variance_text[3])
    ),
    legend = list(title = list(text = "Group"))
  )

# Display the plot
fig
```

## 2D PCA

```{r plotly-2d-pca, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
# Calculate explained variance for PC1 and PC2
explained_variance <- round(100 * (pca$sdev^2 / sum(pca$sdev^2)), 2)
explained_variance_text <- paste0("PC", 1:2, " (", explained_variance[1:2], "%)")

# Extract scores for the first two principal components
pca_data <- as.data.frame(pca$x[, 1:2])  # Select only the first 2 PCs
pca_data$sample <- rownames(pca_data)  # Add sample names as a column
pca_data$group <- samplesheet$group    # Add group information from samplesheet

# Create 2D PCA plot with plotly
fig <- plot_ly(
  data = pca_data,
  x = ~PC1, 
  y = ~PC2,
  color = ~group,
  colors = "Dark2",  # Use a color palette suitable for publication
  text = ~paste("Sample:", sample),  # Add sample names as hover text
  type = "scatter",
  mode = "markers",
  marker = list(size = 12)  # Set marker size for visibility
) %>%
  layout(
    title = "2D PCA Plot of Samples",
    xaxis = list(title = explained_variance_text[1]),
    yaxis = list(title = explained_variance_text[2]),
    legend = list(title = list(text = "Group"))
  )

# Display the plot
fig
```

## Volcano Plot

```{r plotly-volcano, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
# Create a new column for color highlighting based on conditions
res$color <- ifelse(res$gene_symbol %in% highlighted_genes, "orange", 
                    ifelse(res$padj < cutoff_adj_p & abs(res$log2FoldChange) > cutoff_log2fc, "royalblue", "grey"))

# Create a column for label to display the gene symbol only for highlighted genes
res$label <- ifelse(res$gene_symbol %in% highlighted_genes, res$gene_symbol, NA)

# Generate the plotly volcano plot
volcano_plot <- plot_ly(
  data = res,
  x = ~log2FoldChange,
  y = ~-log10(padj),
  text = ~paste("Gene:", gene_symbol, "<br>log2FC:", round(log2FoldChange, 2), "<br>p-adjusted:", signif(padj, 5)),
  color = ~color,
  colors = c("grey" = "grey", "royalblue" = "royalblue", "orange" = "orange"),
  type = "scatter",
  mode = "markers",
  marker = list(size = 5, opacity = 0.7)
) %>%
  layout(
    title = "Volcano Plot",
    xaxis = list(title = "log2 Fold Change"),
    yaxis = list(title = "-log10 Adjusted p-value"),
    showlegend = FALSE
  )

# Add annotations for highlighted genes
volcano_plot <- volcano_plot %>%
  add_trace(
    data = subset(res, !is.na(label)),
    x = ~log2FoldChange,
    y = ~-log10(padj),
    text = ~label,
    mode = "text",
    textposition = "top center",
    showlegend = FALSE
  )

# Display the plot
volcano_plot
```

## MA plot

```{r plotly-maplot, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
# Create a new column for color highlighting based on conditions
res$color <- ifelse(res$gene_symbol %in% highlighted_genes, "orange", 
                    ifelse(res$padj < cutoff_adj_p & abs(res$log2FoldChange) > cutoff_log2fc, "royalblue", "grey"))

# Create a column for label to display the gene symbol only for highlighted genes
res$label <- ifelse(res$gene_symbol %in% highlighted_genes, res$gene_symbol, NA)

# Generate the plotly MA plot
ma_plot <- plot_ly(
  data = res,
  x = ~log2(baseMean + 1),  # Use log2(baseMean + 1) for better visualization on x-axis
  y = ~log2FoldChange,
  text = ~paste("Gene:", gene_symbol, "<br>log2FC:", round(log2FoldChange, 2), "<br>Base Mean:", round(baseMean, 2)),
  color = ~color,
  colors = c("grey" = "grey", "royalblue" = "royalblue", "orange" = "orange"),
  type = "scatter",
  mode = "markers",
  marker = list(size = 5, opacity = 0.7)
) %>%
  layout(
    title = "MA Plot",
    xaxis = list(title = "log2(Base Mean + 1)"),
    yaxis = list(title = "log2 Fold Change"),
    showlegend = FALSE
  )

# Add annotations for highlighted genes
ma_plot <- ma_plot %>%
  add_trace(
    data = subset(res, !is.na(label)),
    x = ~log2(baseMean + 1),
    y = ~log2FoldChange,
    text = ~label,
    mode = "text",
    textposition = "top center",
    showlegend = FALSE
  )

# Display the plot
ma_plot
```

## Heatmap

```{r plotly-heatmap, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
# Select significant genes or top 20 by p-value if no significant genes
significant_genes <- res %>%
  filter(padj < cutoff_adj_p) %>%
  arrange(padj)

if (nrow(significant_genes) == 0) {
  # No significant genes, take top 20 genes with smallest p-values
  significant_genes <- res %>%
    arrange(padj) %>%
    slice_head(n = 20)
}

# Get gene IDs for selected genes
selected_genes <- significant_genes$gene_id

# Extract normalized counts for selected genes
heatmap_data <- normalized_counts[selected_genes, ]
rownames(heatmap_data) <- significant_genes$gene_symbol  # Use gene symbols for row labels

# Perform hierarchical clustering on rows and columns
row_dendrogram <- hclust(dist(heatmap_data))  # Hierarchical clustering for rows
col_dendrogram <- hclust(dist(t(heatmap_data)))  # Hierarchical clustering for columns

# Order rows and columns according to clustering
heatmap_data <- heatmap_data[row_dendrogram$order, col_dendrogram$order]

# Generate the heatmap with plotly
fig <- plot_ly(
  x = colnames(heatmap_data),
  y = rownames(heatmap_data),
  z = log10(as.matrix(heatmap_data) + 1),  # log-transform to improve visualization
  type = "heatmap",
  colorscale = "Viridis",  # Use a professional colormap
  colorbar = list(title = "Expression Level (log10)"),
  showscale = TRUE
) %>%
  layout(
    title = paste0("Heatmap of Significant Genes (N=",length(selected_genes),")"),
    xaxis = list(title = "Samples", tickangle = 45),
    yaxis = list(title = "Genes", tickmode = "array",  showticklabels = FALSE, ticks = "")
  )

# Display the plot
fig
```

# Static figures {.tabset}

## 2D PCA

```{r ggplot-2d-pca, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
# Calculate explained variance for PC1 and PC2
explained_variance <- round(100 * (pca$sdev^2 / sum(pca$sdev^2)), 2)
explained_variance_text <- paste0("PC", 1:2, " (", explained_variance[1:2], "%)")

# Extract scores for the first two principal components
pca_data <- as.data.frame(pca$x[, 1:2])  # Get first two PCs
pca_data$sample <- rownames(pca_data)
pca_data$group <- samplesheet$group  # Add group information from samplesheet

# Create 2D PCA plot with repelled annotations
ggplot(pca_data, aes(x = PC1, y = PC2, color = group, label = sample)) +
  geom_point(size = 3) +  # Set point size for visibility
  geom_text_repel(size = 3, max.overlaps = 10, box.padding = 0.3, point.padding = 0.2) +  # Repelled text labels
  labs(
    title = "2D PCA Plot of Samples",
    x = explained_variance_text[1],
    y = explained_variance_text[2]
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "right"
  ) +
  scale_color_brewer(palette = "Dark2")
```

## Volcano Plot

```{r ggplot-volcano, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
# Create a new column for color highlighting based on conditions
res$color <- ifelse(res$gene_symbol %in% highlighted_genes, "orange", 
                    ifelse(res$padj < cutoff_adj_p & abs(res$log2FoldChange) > cutoff_log2fc, "royalblue", "grey"))
# Create a column for label to display the gene symbol only for highlighted genes
res$label <- ifelse(res$gene_symbol %in% highlighted_genes, res$gene_symbol, NA)
# Generate the ggplot volcano plot
volcano_plot <- ggplot(res, aes(x = log2FoldChange, y = -log10(padj))) +
  geom_point(aes(color = color), size = 2, alpha = 0.7) +
  scale_color_manual(values = c("grey" = "grey", "royalblue" = "royalblue", "orange" = "orange")) +
  labs(
    title = "Volcano Plot",
    x = "log2 Fold Change",
    y = "-log10 Adjusted p-value"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12),
    legend.position = "none"
  )
# Add text labels for highlighted genes
volcano_plot <- volcano_plot +
  geom_text_repel(
    data = subset(res, !is.na(label)),
    aes(label = label),
    size = 3,
    color = "orange",
    max.overlaps = Inf,   # Allow text labels to repel without limiting overlaps
    box.padding = 0.5,    # Add padding around text
    point.padding = 0.3   # Add padding around the points
  )
# Display the plot
print(volcano_plot)
```

## MA plot

```{r ggplot-maplot, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
# Create a new column for color highlighting based on conditions
res$color <- ifelse(res$gene_symbol %in% highlighted_genes, "orange", 
                    ifelse(res$padj < cutoff_adj_p & abs(res$log2FoldChange) > cutoff_log2fc, "royalblue", "grey"))

# Create a column for label to display the gene symbol only for highlighted genes
res$label <- ifelse(res$gene_symbol %in% highlighted_genes, res$gene_symbol, NA)

# Generate the ggplot MA plot with layers in the specified order
ma_plot <- ggplot() +
  # Layer 1: Grey points (non-significant genes)
  geom_point(data = subset(res, color == "grey"), 
             aes(x = log2(baseMean + 1), y = log2FoldChange), 
             color = "grey", size = 2, alpha = 0.5) +
  
  # Layer 2: Blue points (significant genes)
  geom_point(data = subset(res, color == "royalblue"), 
             aes(x = log2(baseMean + 1), y = log2FoldChange), 
             color = "royalblue", size = 2, alpha = 0.7) +
  
  # Layer 3: Orange points (highlighted genes)
  geom_point(data = subset(res, color == "orange"), 
             aes(x = log2(baseMean + 1), y = log2FoldChange), 
             color = "orange", size = 2, alpha = 0.9) +

  # Layer 4: Text labels for highlighted genes
  geom_text_repel(
    data = subset(res, color == "orange" & !is.na(label)),
    aes(x = log2(baseMean + 1), y = log2FoldChange, label = label),
    size = 3,
    color = "orange",
    max.overlaps = Inf,
    box.padding = 0.5,
    point.padding = 0.3
  ) +

  # Axis labels and title
  labs(
    title = "MA Plot",
    x = "log2(Base Mean + 1)",
    y = "log2 Fold Change"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12),
    legend.position = "none"
  )

# Display the plot
print(ma_plot)
```

## Heatmap

```{r ggplot-heatmap, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
# Select significant genes or top 20 by p-value if no significant genes
significant_genes <- res %>%
  filter(padj < cutoff_adj_p) %>%
  arrange(padj)

if (nrow(significant_genes) == 0) {
  # No significant genes, take top 20 genes with smallest p-values
  significant_genes <- res %>%
    arrange(padj) %>%
    slice_head(n = 20)
}

# Get gene IDs for selected genes
selected_genes <- significant_genes$gene_id

# Extract normalized counts for selected genes
heatmap_data <- normalized_counts[selected_genes, ]
rownames(heatmap_data) <- significant_genes$gene_symbol  # Use gene symbols for row labels

# Create a data frame for column annotations
# Assuming `samplesheet` has a column `group` that categorizes samples
column_annotations <- data.frame(Group = samplesheet$group)
rownames(column_annotations) <- samplesheet$sample  # Set row names to match sample names in heatmap_data

# Create the heatmap with hierarchical clustering and column annotations using pheatmap
pheatmap(
  log10(heatmap_data + 1),  # Log-transform for better visualization
  color = colorRampPalette(brewer.pal(9, "YlGnBu"))(100),  # Color scheme
  cluster_rows = TRUE,  # Cluster rows (genes)
  cluster_cols = TRUE,  # Cluster columns (samples)
  scale = "row",  # Standardize each gene's expression across samples
  show_rownames = FALSE,  # Hide row names if they are too many
  show_colnames = TRUE,  # Show column names
  angle_col = 45,  # Rotate x-axis labels by 45 degrees
  fontsize = 10,  # Font size for readability
  main = paste0("Heatmap of Significant Genes (N=",length(selected_genes),")"),
  annotation_col = column_annotations  # Add group annotation to columns
)
```

# Tables {.tabset}

## Statistics

```{r statistics, layout="l-body-outset", echo=TRUE, results='asis', warning=FALSE, message=FALSE}
datatable(
  res,
  extensions = c("FixedColumns", "Buttons"),
  filter = 'top',
  options = list(
    autoWidth = FALSE,
    dom = 'Blftip',
    pageLength = 10,
    searchHighlight = FALSE,
    buttons = c('copy', 'csv', 'excel', 'print'),
    scrollX = TRUE,
    fixedColumns = list(leftColumns = 3),
    order = list(list(7, 'asc'))
  ),
  class = c('compact cell-border stripe hover'),
  rownames = FALSE
) %>%
  formatRound(columns = c(3, 4, 5, 6), digits = 2) %>%  # Round columns 3, 4, 5, and 6 to 2 digits
  formatSignif(columns = c(7, 8), digits = 2)            # Round columns 7 and 8 to 5 digits
```

## Samplesheet

```{r Samplesheet, layout="l-body-outset", echo=FALSE, results='asis', warning=FALSE, message=FALSE}
datatable(
  samplesheet,
  filter = 'top',
  extensions = c("Buttons"),
  options = list(
    autoWidth = FALSE,
    dom = 'Bt',
    searchHighlight = FALSE,
    buttons = c('copy', 'csv', 'excel', 'print'),
    scrollX = TRUE,
    pageLength = -1,        # Show all rows
    paging = FALSE,         # Disable pagination
    columnDefs = list(
      list(className = 'dt-left', targets = "_all")  # Left-align all columns
    )
  ),
  rownames = FALSE
)
```

# Export

```{r export, echo=TRUE, results='markup', warning=FALSE, message=TRUE}
export_filename <- paste0("DGEA_", params$filetag, "_export.xlsx")
# Create a new workbook
wb <- createWorkbook()
# Add each data frame as a new sheet
addWorksheet(wb, "Samplesheets")
writeData(wb, "Samplesheets", samplesheet)
addWorksheet(wb, "Statistics")
writeData(wb, "Statistics", res, row.names = FALSE)
addWorksheet(wb, "Norm. counts")
writeData(wb, "Norm. counts", out_norm, row.names = TRUE)
# Save the workbook
saveWorkbook(wb, export_filename, overwrite = TRUE)
export_csv <- paste0("DGEA_", params$filetag, "_stat.csv")
write.csv(res, export_csv, row.names = FALSE)
```

- Please download the export excel file: [`r export_filename`](`r export_filename`)
- Or download the CSV file: [`r export_filename`](`r export_filename`)
- If you desire to annotate the volcano plot and annotate the genes you prefer, please try out our web app [AnnoFig](https://genomics.rwth-aachen.de/annofig/) with the CSV file above.

# References

<div id="refs"></div>

# R session information

```{r DGEA_FILETAG-session, echo=FALSE, results='markup', warning=FALSE, message=TRUE}
sessionInfo()
```