---
title: "GPM_PROJECTNAME"
author:
RMD_AUTHORS
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: false
    theme: lumen
bibliography: references.bib
---

```{r conig, echo=FALSE, results="hide", warning=FALSE, message=FALSE}
library(stringr)
library("readr")
# library(tidyr)
# library(data.table)
options(warn = -1)
options(knitr.duplicate.label = "allow")

################# File paths #####################################################
DIR_base <- "PROJECT_PROJECT_PATH/"
DIR_project <- paste0(DIR_base, "analysis/")
DIR_cellranger <- paste0(DIR_base, "cellranger/")
DIR_multi_output <- paste0(DIR_cellranger, "multi_output/")
DIR_multi_output_published <- "EXPORT_URL/PROJECT_PROJECT_NAME/2_Processed_data/multi_output/"
```

![](RMD_INSTITUTE_LOGO)

This document is the summary of the bioinformatic analysis done by Genomic Facility in IZKF, RWTH Aachen Uniklinik. Please contact us for any further discussion or collaboration. If these results are used in any poster or publications, please acknowledge our core facility with the text below:

> This work was supported by the Genomics Facility, a core facility of the Interdisciplinary Center for Clinical Research (IZKF) Aachen within the Faculty of Medicine at RWTH Aachen University.

Please refer to our [Publication Policy](https://genomics.rwth-aachen.de/services/publication-policy/).

***

# Download/Browse the files

1. [Raw Data](EXPORT_URL/PROJECT_PROJECT_NAME/1_Raw_data): 
    * BCL files from the sequencer
    * FASTQ files after demultiplexing
2. [Processed Data](EXPORT_URL/PROJECT_PROJECT_NAME/2_Processed_data): 
    * summary for the results
    * count folder
      + web summary file
      + cloupe file for browsing in Loupe browser
      + h5 files for downstream analysis
      + sample_feature_bc_matrix
      + bam and bai files
      + other analysis results (umap, tsne, pca, diffexp and clustering)
    * vdj folder
      + vloupe file for browsing in Loupe browser
      + consensus files (fasta, fai, bam, bai, annotation)
      + filtered contig (fasta, annotation)
      + concat_ref (fasta, fai, bam, bai)
      + High-level description of each clonotype
      + List of barcodes identified as T/B cells
      + Annotated contigs and consensus sequences of V(D)J rearrangements in the AIRR format
3. [Reports](EXPORT_URL/PROJECT_PROJECT_NAME/3_Reports):
    * [MultiQC reports](EXPORT_URL/PROJECT_PROJECT_NAME/3_Reports/multiqc/multiqc_report.html) for the quality check
    * Basic analysis of the Genomics Facility (This report)
4. [Compressed files for download](GPM_URL_TAR):
    * Three tar files are available for download the above three folders.
  
In the command line environment, you can also download everything by the commands below with the given username and password from us:

```
wget -r -nH -np -nd -P --user USERNAME --password PASSWORD -A '*.tar' EXPORT_URL/PROJECT_PROJECT_NAME/compressed_tar/ -P PROJECT_PROJECT_NAME
```

Please make sure that you store these data in a lab-wise level. Please refer to our [Data Management Policy](https://genomics.rwth-aachen.de/services/data-management-policy/).

\textcolor{red}{Your data will be deleted according to this policy automatically. Please download everything.}

# Method description

Raw data was demultiplexed with cellranger mkfastq (10x Genomics v7.1.0) to generate fastq files for both GEX and VDJ libraries. The fastq files were then analyzed by the cellranger multi with default parameters. This report was generated using custom scripts in R version 4.2.2. 

# Reports of Cellranger multi for GEX and VDJ

```{r load, echo=FALSE, results="hide", warning=FALSE, message=FALSE}
samples <- list.dirs(path = DIR_multi_output, full.names = F, recursive = F)
summaryFiles <- vector(mode="list", length=length(samples))
names(summaryFiles) <- samples
for (i in 1:length(samples)) {
  summaryFiles[[i]] <- paste0(DIR_multi_output_published,samples[[i]],"/outs/web_summary.html")
}
```

```{r, echo=FALSE, results="asis"}
for( i in 1:length(samples)){
  cat("\n")
  cat("## ", "[",samples[[i]],"]","(",summaryFiles[[i]],")\n", sep="") # Create links to the
  cat("\n")
}
```


# References

<div id="refs"></div>

# R session information

```{r info, echo=FALSE, results='markup', warning=FALSE, message=TRUE}
sessionInfo()
```