---
title: "PROJECT_PROJECT_STRING"
author:
RMD_AUTHORS
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: false
    theme: lumen
bibliography: references.bib
---

```{r conig, echo=FALSE, results="hide", warning=FALSE, message=FALSE}
library("heatmaply")
library("readr")
require("tidyr")
options(warn = -1)

DIR_base <- "PROJECT_PROJECT_PATH/"
```

![](RMD_INSTITUTE_LOGO)

This document is the summary of the basic bacterial genome assembly done by Genomic Core Facility in IZKF, RWTH Aachen Uniklinik. Please contact us for any further discussion or collaboration.

Please refer to our [Publication Policy](https://genomics.rwth-aachen.de/services/publication-policy/).

------------------------------------------------------------------------

The following contents are provided after the basic assembly analysis of the bacterial whole genome sequencing data with Bacass pipeline[@Peltzer2021-xn] under nf-core framework[@Ewels2020-il] :

- Quality reports of the sequencing data
- Contamination check reports for the samples
- Assenbly results
- Assembly quality assessment reports for the samples
- Annotation of the assembled genome

------------------------------------------------------------------------

# Quality report of the sequencing data

The raw sequencing data were demultiplexed. QC assessment was applied on the resulted fastq files. Bellow is the resulted multiQC[@Ewels2016-pr] report:

[MultiQC reports](EXPORT_URL/PROJECT_PROJECT_NAME/3_Reports/multiqc/multiqc_report.html)

# Contamination check

The raw data from a genome sequencing could contain DNA from contaminating organisms. In this step, sequencing results were scanned using metagenomic analysis tool Kraken2[@Wood2019-er] to identify potential contamination.

[Contamination reports](EXPORT_URL/PROJECT_PROJECT_NAME/2_Processed_data/Kraken2)

# Genome Assembly

Unicycler[@Wick2017-us] was applied to assemble the short sequencing reads. The assembled scaffolds fasta files are listed bellow:

[Assembled scaffolds](EXPORT_URL/PROJECT_PROJECT_NAME/2_Processed_data/assembly)

# Assembly quality assessment

The quality of the assembled contigs were evaluated by QUAST[@Gurevich2013-tw] with the following report:

[Assembly quality report](EXPORT_URL/PROJECT_PROJECT_NAME/2_Processed_data/QUAST/other_files/report.html)

# Annotation the assembled genome

The assemblies are annotated by Prokka[@Seemann2014-ic].

[Assembly quality report](EXPORT_URL/PROJECT_PROJECT_NAME/2_Processed_data/Prokka)

# References

<div id="refs"></div>


# R session information

```{r info, echo=FALSE, results='markup', warning=FALSE, message=TRUE}
sessionInfo()
```